---
title: "ECHO CNV Global Report"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
params:
  cnv_calls: NA
  qc_metrics: NA
  counts: NA
  bed_file: NA
  models: NA
  refs: NA
  settings: NA
  config: NA
  sample_table: NA
  ref_bams: NA
  log_file: NA
  pdf_output: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

library(DT)
library(plotly)
library(ggplot2)
library(dplyr)
library(knitr)
library(tidyr)
library(htmltools)
library(ExomeDepth)
library(kableExtra)

# Helper function to get CNV-specific data for plotting
get_cnv_plot_data <- function(call_row, counts, bed_file, models, refs, gene_gap = 1) {

  sample <- call_row$Sample
  idx_start <- if ("global_start" %in% colnames(call_row)) as.numeric(call_row$global_start) else as.numeric(call_row$Start.p)
  idx_end <- if ("global_end" %in% colnames(call_row)) as.numeric(call_row$global_end) else as.numeric(call_row$End.p)

  ref_samples <- refs[[sample]]

  if (is.null(ref_samples) || length(ref_samples) == 0) {
    return(NULL)
  }

  exon_range <- seq(
    max(1, idx_start - 5),
    min(nrow(bed_file), idx_end + 5)
  )

  if (length(exon_range) == 0) {
    return(NULL)
  }

  # Gap-inserted x-axis positions
  gap_pos    <- compute_gene_gap_positions(bed_file, exon_range, gap = gene_gap)
  px         <- gap_pos$px
  gene_group <- gap_pos$gene_group

  test_counts <- counts[exon_range, sample]

  ref_counts <- rowSums(
    counts[exon_range, ref_samples, drop = FALSE]
  )

  test_median <- median(test_counts)
  ref_median <- median(ref_counts)

  expected <- ref_counts * (test_median / ref_median)
  expected_safe <- pmax(expected, 1)

  totals <- test_counts + ref_counts

  p_expected <- expected_safe / pmax(totals, 1)
  p_expected <- pmin(pmax(p_expected, 1e-6), 1 - 1e-6)

  ratio <- test_counts / expected_safe

  rho <- max(models[[sample]][1], 0.005)

  ci_low <- vapply(seq_along(exon_range), function(i) {
    ECHO:::qbetabinom(
      0.025,
      totals[i],
      rho,
      p_expected[i]
    ) / expected_safe[i]
  }, numeric(1))

  ci_high <- vapply(seq_along(exon_range), function(i) {
    ECHO:::qbetabinom(
      0.975,
      totals[i],
      rho,
      p_expected[i]
    ) / expected_safe[i]
  }, numeric(1))

  # Z-score data
  var_betabinom <- totals * p_expected * (1 - p_expected) * (1 + (totals - 1) * rho)
  var_betabinom <- pmax(var_betabinom, 1)
  sd_z <- sqrt(var_betabinom)
  test_z <- (test_counts - expected_safe) / sd_z

  ref_z_mat <- sapply(ref_samples, function(r) {
    scaling_r    <- test_median / median(counts[exon_range, r])
    ref_r_scaled <- counts[exon_range, r] * scaling_r
    (ref_r_scaled - expected_safe) / sd_z
  })
  if (is.null(dim(ref_z_mat))) ref_z_mat <- matrix(ref_z_mat, ncol = length(ref_samples))
  colnames(ref_z_mat) <- ref_samples

  exon_labels <- if ("exon_number" %in% colnames(bed_file) && any(!is.na(bed_file$exon_number))) {
    bed_file$exon_number[exon_range]
  } else if ("exon" %in% colnames(bed_file) && any(!is.na(bed_file$exon))) {
    bed_file$exon[exon_range]
  } else {
    ave(seq_len(nrow(bed_file)),
        bed_file$gene,
        FUN = seq_along)[exon_range]
  }

  # ---- ADD Gene column ----
  gene_col <- if ("gene" %in% colnames(bed_file)) "gene" else "Gene"
  gene_names <- bed_file[[gene_col]][exon_range]
  gene_names[is.na(gene_names)] <- ""

  affected <- exon_range %in% (idx_start:idx_end)
  bg_calib <- check_background_calibration(ratio, ci_low, ci_high, affected)

  df <- data.frame(
    Exon = exon_range,
    ExonLabel = exon_labels,
    Gene = gene_names,        # <-- ADDED
    Px = px,
    GeneGroup = gene_group,
    
    Observed = test_counts,
    Expected = expected_safe,
    Ratio = ratio,
    CI_low = ci_low,
    CI_high = ci_high,
    Z = test_z,
    Affected = affected
  )
  attr(df, "ref_z_mat") <- ref_z_mat
  attr(df, "bg_calibration") <- bg_calib
  df
}
```

<style>
  .alert-warning { border-left: 5px solid #f0ad4e; margin-bottom: 20px; background-color: #fcf8e3; color: #8a6d3b; padding: 15px; border-radius: 4px; }
  .dt-buttons { margin-bottom: 15px; }
  .btn-default { background-color: #34495e !important; color: white !important; border: none !important; }
  .btn-default:hover { background-color: #2c3e50 !important; }
</style>

## Summary

- **Number of samples with CNV calls**:
  `r length(unique(params$cnv_calls$Sample))`

- **Total CNV calls**:
  `r nrow(params$cnv_calls)`

- **Report generated**:
  `r Sys.time()`

## QC Metrics Overview (all samples)

```{r qc_overview, results='asis'}
# 1. Output Pipeline Alerts for missing chromosomes if present
if (!is.null(params$qc_metrics) && is.data.frame(params$qc_metrics) && nrow(params$qc_metrics) > 0) {
  missing_chrs <- params$qc_metrics[params$qc_metrics$Type == "Missing Chromosome", ]
  if (nrow(missing_chrs) > 0) {
    for (i in seq_len(nrow(missing_chrs))) {
      # BUG FIX: this used to be cat(sprintf('\n      <div ...')) with each
      # HTML line indented to match the surrounding R code. results='asis'
      # output goes straight into the Markdown source Pandoc then parses,
      # and Pandoc's Markdown reader treats a line indented 4+ spaces as an
      # INDENTED CODE BLOCK, not raw HTML — so this rendered as literal,
      # HTML-escaped text in a <pre><code> block instead of a styled alert
      # box. Fixed by emitting each line via its own cat() call starting at
      # column 0 (matching the working style already used successfully
      # elsewhere in this file, e.g. the confidence_recap chunk below).
      cat('<div class="alert alert-warning" role="alert">\n')
      cat('<h4 class="alert-heading" style="margin-top:0;font-weight:bold;"><i class="fa fa-exclamation-triangle"></i> Pipeline Warning</h4>\n')
      cat(sprintf('<p style="margin-bottom:0;">%s. <strong>Analysis for this chromosome has been skipped automatically.</strong></p>\n',
                  htmltools::htmlEscape(missing_chrs$Details[i])))
      cat('</div>\n')
    }
    cat("<br>")
  }
}

# 2. Render main table excluding technical missing chromosome warnings
if (!is.null(params$qc_metrics) && is.data.frame(params$qc_metrics) && nrow(params$qc_metrics) > 0) {
  display_qc <- params$qc_metrics[params$qc_metrics$Type != "Missing Chromosome", ]
  
  if (nrow(display_qc) > 0) {
    DT::datatable(
      display_qc,
      extensions = 'Buttons',
      options = list(
        pageLength = 15,
        scrollX = TRUE,
        dom = 'Bfrtip',
        buttons = list(
          list(
            extend = 'colvis',
            text = '<i class="fa fa-eye"></i> Show/Hide Columns',
            collectionLayout = 'fixed columns',
            fade = 100
          ),
          'copy', 'csv'
        )
      ),
      caption = "QC flags across all samples"
    )
  } else {
    cat("<p>No sample-specific QC warnings or issues recorded.</p>\n")
  }
} else {
  cat("<p><em>No QC metrics file supplied — QC section skipped.</em></p>\n")
}
```

---

## Principal Component Analysis (PCA) of Coverage Profiles

```{r pca_coverage}
counts <- params$counts
if (!is.null(counts) && is.data.frame(counts) && ncol(counts) >= 3) {
  # Identify sample columns (exclude metadata)
  metadata_cols <- c("chromosome", "start", "end", "exon", "exon_number", "GC", "gene")
  sample_cols <- setdiff(colnames(counts), metadata_cols)
  
  if (length(sample_cols) >= 3) {
    count_mat <- as.matrix(counts[, sample_cols, drop = FALSE])
    # Log2 transform and remove zero-variance rows
    log_mat <- log2(count_mat + 1)
    row_var <- apply(log_mat, 1, var, na.rm = TRUE)
    keep <- row_var > 0 & !is.na(row_var)
    if (sum(keep) >= 2) {
      log_mat <- log_mat[keep, ]
      pca <- prcomp(t(log_mat), scale. = TRUE, center = TRUE)
      var_exp <- summary(pca)$importance[2, ] * 100
      pca_df <- data.frame(PC1 = pca$x[,1], PC2 = pca$x[,2], Sample = sample_cols)
      
      # Add gender information if sample_table is provided
      sample_table <- params$sample_table
      if (!is.null(sample_table) && is.data.frame(sample_table) && 
          "gender" %in% colnames(sample_table)) {
        # Ensure column name matches (sample_name or sample)
        if ("sample_name" %in% colnames(sample_table)) {
          pca_df <- merge(pca_df, sample_table[, c("sample_name", "gender")], 
                          by.x = "Sample", by.y = "sample_name", all.x = TRUE)
        } else if ("Sample" %in% colnames(sample_table)) {
          pca_df <- merge(pca_df, sample_table[, c("Sample", "gender")], 
                          by = "Sample", all.x = TRUE)
        } else {
          pca_df$Group <- "All samples"
        }
        color_by <- if ("gender" %in% colnames(pca_df)) "gender" else "Group"
      } else {
        pca_df$Group <- "All samples"
        color_by <- "Group"
      }
      
      # Interactive plotly scatter plot
      p <- plot_ly(pca_df, x = ~PC1, y = ~PC2, text = ~Sample,
                   color = as.formula(paste0("~", color_by)),
                   type = 'scatter', mode = 'markers+text',
                   textposition = 'top right',
                   marker = list(size = 10),
                   hovertemplate = paste('<b>%{text}</b><br>PC1: %{x:.2f}<br>PC2: %{y:.2f}<extra></extra>')) %>%
        layout(title = paste("PCA of Coverage Profiles (", ncol(count_mat), " samples, ", nrow(log_mat), " exons)"),
               xaxis = list(title = paste0("PC1 (", round(var_exp[1], 1), "%)")),
               yaxis = list(title = paste0("PC2 (", round(var_exp[2], 1), "%)")),
               legend = list(title = list(text = ifelse(color_by == "gender", "Gender", "Group"))))
      p
    } else {
      cat("Insufficient variation for PCA after filtering (less than 2 variable exons).")
    }
  } else {
    cat("Not enough samples (≥3 required) for PCA.")
  }
} else {
  cat("Counts data not available or insufficient dimensions for PCA.")
}
```

# Per-sample Reports

```{r per_sample_loop, results='asis'}

samples <- unique(params$cnv_calls$Sample)

if (length(samples) == 0) {

  cat("No CNV calls found in any sample.\n")

} else {

  for (s in samples) {

    cat("\n\n## ", s, "\n")

    # QC metrics for this sample
    if (!is.null(params$qc_metrics) && is.data.frame(params$qc_metrics)) {
      qc_s <- params$qc_metrics[
        (params$qc_metrics$Sample == s | params$qc_metrics$Sample == "All") & 
         params$qc_metrics$Type != "Missing Chromosome",
      ]

      if (nrow(qc_s) > 0) {

        cat("\n### QC flags\n")

        w_qc <- DT::datatable(
          qc_s,
          options = list(
            dom = 't',
            pageLength = 5
          )
        )

        rendered_qc <- htmltools::renderTags(w_qc)

        knitr::knit_meta_add(
          rendered_qc$dependencies
        )

        cat(rendered_qc$html)
        cat("\n")
      }
    }

    # Key metrics for this sample
    if (!is.null(params$counts) &&
        s %in% colnames(params$counts)) {

      cov <- params$counts[, s]

      median_depth <- median(
        cov,
        na.rm = TRUE
      )

      total_reads <- sum(
        cov,
        na.rm = TRUE
      )

      cat("\n### Key metrics\n")

      cat(
        paste0(
          "- Median exon depth: ",
          round(median_depth, 1),
          "\n"
        )
      )

      cat(
        paste0(
          "- Total reads: ",
          format(total_reads, big.mark = ","),
          "\n"
        )
      )

      if (!is.null(params$settings$min_cov) &&
          median_depth < params$settings$min_cov) {

        cat(
          "⚠️ **Below minimum depth threshold**\n"
        )
      }

      if (!is.null(params$settings$min_total_reads) &&
          total_reads < params$settings$min_total_reads) {

        cat(
          "⚠️ **Below minimum total reads**\n"
        )
      }
    }

    # CNV table for this sample
    cnv_s <- params$cnv_calls[
      params$cnv_calls$Sample == s,
    ]

    cat("\n### CNV calls\n")

    if (nrow(cnv_s) > 0) {

      cnv_display <- cnv_s[, c("Sample", "Chromosome", "Gene", "Start", "End",
                               "Type", "Nexons", "Reads.ratio", "Confidence")]
      cnv_display$Type <- ifelse(cnv_display$Type == "duplication", "DUP",
                           ifelse(cnv_display$Type == "deletion", "DEL",
                                  toupper(cnv_display$Type)))
      cnv_display$Reads.ratio <- round(cnv_display$Reads.ratio, 2)
      rownames(cnv_display) <- NULL

      w_cnv <- DT::datatable(
        cnv_display,
        rownames = FALSE,
        colnames = c("Sample", "Chr", "Gene", "Start", "End",
                     "Type", "Number of exons", "Fold change", "Confidence"),
        extensions = "Buttons",
        options = list(
          dom = "Bfrtip",
          buttons = list(
            list(
              extend = 'colvis',
              text = '<i class="fa fa-eye"></i> Show/Hide Columns',
              collectionLayout = 'fixed columns',
              fade = 100
            ),
            "copy", "csv", "excel"
          ),
          autoWidth = FALSE,
          columnDefs = list(
            list(width = "90px",  targets = 0),  # Sample
            list(width = "55px",  targets = 1),  # Chr
            list(width = "90px",  targets = 2),  # Gene
            list(width = "80px",  targets = 3),  # Start
            list(width = "80px",  targets = 4),  # End
            list(width = "45px",  targets = 5),  # Type
            list(width = "50px",  targets = 6),  # Exons
            list(width = "80px",  targets = 7),  # Fold change
            list(width = "90px",  className = "dt-center", targets = 8)  # Confidence
          )
        ),
        caption = paste("Detected CNVs for", s),
        width = "100%"
      ) %>%
        formatStyle(
          "Confidence",
          backgroundColor = styleEqual(
            c("HIGH", "MEDIUM", "LOW"),
            c("lightgreen", "yellow", "salmon")
          )
        )

      rendered_cnv <- htmltools::renderTags(
        w_cnv
      )

      knitr::knit_meta_add(
        rendered_cnv$dependencies
      )

      cat(rendered_cnv$html)
      cat("\n")

    } else {

      cat(
        "No CNV calls detected for this sample.\n"
      )
    }

    # Warning for low-confidence genes
    low_conf_genes <- params$settings$score_low_confidence_genes

    all_genes <- unique(
      unlist(
        strsplit(cnv_s$Gene, "_")
      )
    )

    flag <- intersect(
      all_genes,
      low_conf_genes
    )

    if (length(flag) > 0) {

      cat(
        "\n⚠️ **Warning:** This sample has CNVs involving genes with known pseudogene issues: ",
        paste(flag, collapse = ", "),
        ". Consider orthogonal validation.\n"
      )
    }

    # Interactive plots for each CNV
    if (nrow(cnv_s) > 0) {

      cat("\n### CNV plots\n")

      for (i in seq_len(nrow(cnv_s))) {

        call <- cnv_s[i, ]

        cat(
          "\n#### ",
          call$Gene,
          " (",
          ifelse(call$Type == "duplication", "DUP", ifelse(call$Type == "deletion", "DEL", toupper(call$Type))),
          ")\n"
        )

        cat(
          "*Confidence: ",
          call$Confidence,
          " | Fold change: ",
          round(call$Reads.ratio, 2),
          "*\n\n"
        )

        # Prepare CNV plotting data using plots.R logic
        plot_data <- get_cnv_plot_data(
          call,
          params$counts,
          params$bed_file,
          params$models,
          params$refs,
          gene_gap = params$settings$plot_gene_gap %||% 1
        )

        if (is.null(plot_data) || nrow(plot_data) == 0) {
          cat("*(Insufficient data to generate plot)*\n")
          next
        }

        # New: per-call diagnostic (ported from CANOPE, see
        # check_background_calibration() in utils.R). Informational only —
        # doesn't change the call, the interval, or the confidence score.
        bg_calib <- attr(plot_data, "bg_calibration")
        if (isTRUE(bg_calib$flag)) {
          cat("\n> \u26a0\ufe0f **Background calibration flag:** ",
              round(bg_calib$pct_outside), "% of non-called exons in this window (",
              bg_calib$n_outside, "/", bg_calib$n_background,
              ") fall outside the modeled 95% interval \u2014 well above the ~5% expected under a ",
              "well-calibrated model. This is a diagnostic flag, not a correction: it can mean the real ",
              "alteration extends beyond the exons called here, the reference set is a weaker match for ",
              "this sample/region, or there is a technical/batch difference at this locus. Worth a manual ",
              "look before reporting this call with full confidence.\n\n", sep = "")
        }

        ratio_subtitle <- if (isTRUE(bg_calib$flag)) {
          paste0(round(bg_calib$pct_outside), "% of background exons outside CI (",
                bg_calib$n_outside, "/", bg_calib$n_background, ") \u2014 check region/reference match")
        } else NULL

        # Ratio plot with CI ribbon around reference expectation
        p_ratio <- ggplot(plot_data, aes(x = Px)) +
          geom_ribbon(
            aes(
              ymin = CI_low,
              ymax = CI_high,
              group = GeneGroup
            ),
            fill = "grey80",
            alpha = 0.35
          ) +
          geom_hline(
            yintercept = 1,
            linetype = "dashed",
            colour = "grey40"
          ) +
          geom_point(
            data = subset(plot_data, !Affected),
            aes(
              y = Ratio,
              color = "Reference samples",
              text = paste0("Sample: ", call$Sample, "<br>Exon: ", ExonLabel,
                            "<br>Gene: ", Gene,   # <-- ADDED
                            "<br>Observed/Expected: ", round(Ratio, 3)
              )
            ),
            size = 3
          ) +
          geom_point(
            data = subset(plot_data, Affected),
            aes(
              y = Ratio,
              color = "Affected exon(s)",
              text = paste0("Sample: ", call$Sample, "<br>Exon: ", ExonLabel,
                            "<br>Gene: ", Gene,   # <-- ADDED
                            "<br>Observed/Expected: ", round(Ratio, 3),
                            "<br>Confidence: ", call$Confidence
              )
            ),
            size = 3.5
          ) +

          labs(
            y = "Observed / Expected",
            x = NULL,
            subtitle = ratio_subtitle
          ) +
          theme_bw() +
          scale_color_manual(
            values = c(
              "Reference samples" = "steelblue",
              "Affected exon(s)" = "red"
            ),
            name = NULL
          ) +
          theme(
            legend.position = "top",
            axis.text.x = element_text(
              angle = 45,
              size = 8,
              hjust = 1
            )
          ) +
          scale_x_continuous(
            breaks = plot_data$Px,
            labels = plot_data$ExonLabel
          )  

        # Coverage plot aligned with plots.R lower panel
        ref_samples_s <- params$refs[[s]]
        exon_range_s  = plot_data$Exon
        test_median_s <- median(params$counts[exon_range_s, s])

        ref_rows <- lapply(ref_samples_s, function(r) {
          scaling <- test_median_s / median(params$counts[exon_range_s, r])
          data.frame(
            Exon      = exon_range_s,
            Px        = plot_data$Px,
            Count     = log(pmax(params$counts[exon_range_s, r] * scaling, 1)),
            Group     = "Reference samples",
            ExonLabel = plot_data$ExonLabel,
            Gene      = plot_data$Gene,   # <-- ADDED
            SampleID  = r,
            stringsAsFactors = FALSE
          )
        })

        obs_df <- data.frame(
          Exon      = plot_data$Exon,
          Px        = plot_data$Px,
          Count     = log(pmax(plot_data$Observed, 1)),
          Group     = ifelse(plot_data$Affected, "Affected exon(s)", "Test sample"),
          ExonLabel = plot_data$ExonLabel,
          Gene      = plot_data$Gene,   # <-- ADDED
          SampleID  = call$Sample,
          stringsAsFactors = FALSE
        )

        p_cov_data <- bind_rows(c(ref_rows, list(obs_df)))

        p_cov <- ggplot() +
          geom_point(
            data = subset(p_cov_data, Group == "Reference samples"),
            aes(
              x = Px,
              y = Count,
              color = Group,
              text = paste0("Sample: ", SampleID, "<br>Exon: ", ExonLabel,
                "<br>Log coverage: ", round(Count, 3)
              )
            ),
            alpha = 0.7,
            size = 2.5
          ) +
          geom_point(
            data = subset(p_cov_data, Group == "Test sample"),
            aes(
              x = Px,
              y = Count,
              color = Group,
              text = paste0("Sample: ", SampleID, "<br>Exon: ", ExonLabel,
                "<br>Log coverage: ", round(Count, 3)
              )
            ),
            size = 2.5
          ) +
          geom_point(
            data = subset(p_cov_data, Group == "Affected exon(s)"),
            aes(
              x = Px,
              y = Count,
              color = Group,
              text = paste0("Sample: ", SampleID, "<br>Exon: ", ExonLabel,
                "<br>Log coverage: ", round(Count, 3),
                "<br>Confidence: ", call$Confidence
              )
            ),
            size = 3.5
          ) +
          labs(
            y = "Log(Coverage)",
            x = NULL
          ) +
          theme_bw() +
          scale_color_manual(
            values = c(
              "Reference samples" = "grey50",
              "Test sample" = "blue",
              "Affected exon(s)" = "red"
            ),
            name = NULL
          ) +
          theme(
            legend.position = "top",
            axis.text.x = element_text(
              angle = 45,
              size = 8,
              hjust = 1
            )
          ) +
          scale_x_continuous(
            breaks = plot_data$Px,
            labels = plot_data$ExonLabel
          ) 

        # Z-score panel: test sample vs. each individual reference sample,
        # all in units of the beta-binomial model's own implied SD (see
        # get_cnv_plot_data() above). Ported from CANOPE.
        ref_z_mat <- attr(plot_data, "ref_z_mat")
        ref_z_rows <- lapply(colnames(ref_z_mat), function(r) {
          data.frame(
            Exon      = plot_data$Exon,
            Px        = plot_data$Px,
            GeneGroup = plot_data$GeneGroup,
            Z         = ref_z_mat[, r],
            Group     = "Reference samples",
            ExonLabel = plot_data$ExonLabel,
            RefName   = r,
            stringsAsFactors = FALSE
          )
        })
        z_obs_df <- data.frame(
          Exon      = plot_data$Exon,
          Px        = plot_data$Px,
          GeneGroup = plot_data$GeneGroup,
          Z         = plot_data$Z,
          Group     = ifelse(plot_data$Affected, "Affected exon(s)", "Test sample"),
          ExonLabel = plot_data$ExonLabel,
          RefName   = call$Sample,
          stringsAsFactors = FALSE
        )
        p_z_data <- bind_rows(c(ref_z_rows, list(z_obs_df)))

        p_zscore <- ggplot() +
          geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
          geom_point(
            data = subset(p_z_data, Group == "Reference samples"),
            aes(x = Px, y = Z, group = RefName,
                text = paste0("Sample: ", RefName, "<br>Exon: ", ExonLabel,
                              "<br>Gene: ", Gene,   # <-- ADDED
                              "<br>Z-score: ", round(Z, 2))),
            colour = "grey60", size = 1.6, alpha = 0.8
          ) +
          geom_point(
            data = subset(p_z_data, Group == "Test sample"),
            aes(x = Px, y = Z, color = Group,
                text = paste0("Sample: ", RefName, "<br>Exon: ", ExonLabel,
                              "<br>Gene: ", Gene,   # <-- ADDED
                              "<br>Z-score: ", round(Z, 2))),
            size = 3
          ) +
          geom_point(
            data = subset(p_z_data, Group == "Affected exon(s)"),
            aes(x = Px, y = Z, color = Group,
                text = paste0("Sample: ", RefName, "<br>Exon: ", ExonLabel,
                              "<br>Gene: ", Gene,   # <-- ADDED
                              "<br>Z-score: ", round(Z, 2),
                              "<br>Confidence: ", call$Confidence)),
            size = 3.5
          ) +
          labs(y = "Z-score vs references", x = NULL) +
          theme_bw() +
          scale_color_manual(
            values = c("Test sample" = "blue", "Affected exon(s)" = "red"),   # <-- CHANGED
            name = NULL
          ) +
          labs(y = "Z-score vs references", x = NULL) +
          theme_bw() +
          scale_color_manual(
            values = c("Test sample" = "red", "Affected exon(s)" = "darkred"),
            name = NULL
          ) +
          theme(
            legend.position = "top",
            axis.text.x = element_text(angle = 45, size = 8, hjust = 1)
          ) +
          scale_x_continuous(
            breaks = plot_data$Px,
            labels = plot_data$ExonLabel
          )

        # Convert to interactive and adjust layout to prevent truncation
        w_ratio <- ggplotly(p_ratio, tooltip = "text", height = 400) %>%
          layout(
                margin = list(t = 60, b = 80),
                legend = list(orientation = "h", y = 1.1))

        w_cov <- ggplotly(p_cov, tooltip = "text", height = 400) %>%
          layout(
                margin = list(t = 60, b = 80),
                legend = list(orientation = "h", y = 1.1))

        w_zscore <- ggplotly(p_zscore, tooltip = "text", height = 400) %>%
          layout(
                margin = list(t = 60, b = 80),
                legend = list(orientation = "h", y = 1.1))

        # Render with proper dependency handling
        rendered_ratio <- htmltools::renderTags(w_ratio)
        knitr::knit_meta_add(rendered_ratio$dependencies)
        cat(rendered_ratio$html)
        cat("\n")

        rendered_cov <- htmltools::renderTags(w_cov)
        knitr::knit_meta_add(rendered_cov$dependencies)
        cat(rendered_cov$html)
        cat("\n")

        rendered_zscore <- htmltools::renderTags(w_zscore)
        knitr::knit_meta_add(rendered_zscore$dependencies)
        cat(rendered_zscore$html)
        cat("\n")
      }
    }
  }
}
```

```{r pdf_output, include=FALSE}
if (isTRUE(params$pdf_output) || identical(params$pdf_output, "true")) {
  if (!requireNamespace("pagedown", quietly = TRUE)) {
    warning("Package 'pagedown' is needed for PDF output but is not installed.")
  } else {
    pdf_out <- sub("\\.html$", ".pdf", knitr::current_input(dir = TRUE))
    pagedown::chrome_print(knitr::current_input(dir = TRUE), output = pdf_out)
    message("[INFO] PDF saved to: ", pdf_out)
  }
}
```

---
## Confidence Score Method

<details>
<summary><strong>&#9432; Click to expand/collapse confidence scoring rules</strong></summary>

```{r confidence_recap, results='asis'}
settings <- params$settings

# Define the three categories with their criteria
score_matrix <- data.frame(
  Category = c("**HIGH**", "**MEDIUM**", "**LOW**"),
  Correlation = c(
    sprintf("≥ %.3f", settings$score_high_corr %||% 0.985),
    sprintf("≥ %.3f", settings$score_med_corr %||% 0.95),
    "Any"
  ),
  `References` = c(
    sprintf("≥ %d", settings$score_high_refs %||% 3),
    sprintf("≥ %d", settings$score_med_refs %||% 2),
    "Any"
  ),
  `Fold change` = c(
    sprintf("≤ %.2f **or** ≥ %.2f", settings$score_high_ratio_low %||% 0.70, settings$score_high_ratio_high %||% 1.30),
    sprintf("≤ %.2f **or** ≥ %.2f", settings$score_low_ratio_low %||% 0.75, settings$score_low_ratio_high %||% 1.25),
    "Intermediate"
  ),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

forced_genes <- settings$score_low_confidence_genes
if (is.null(forced_genes)) forced_genes <- c()

cat('<div style="margin: 10px 0 20px 0;">\n')
cat('<p><strong>Assignment rules:</strong></p>\n')
cat('<ul>\n')
cat('<li><strong>HIGH</strong> only if <em>all</em> HIGH criteria are met.</li>\n')
cat('<li><strong>MEDIUM</strong> if HIGH not met but <em>all</em> MEDIUM criteria are met.</li>\n')
cat('<li><strong>LOW</strong> otherwise.</li>\n')
if (length(forced_genes) > 0) {
  cat(sprintf('<li><strong>Always LOW</strong> (genes with homology issues): <em>%s</em>.</li>\n',
              paste(forced_genes, collapse = ", ")))
}
cat('</ul>\n')
cat('</div>\n')

knitr::kable(score_matrix, format = "html", escape = FALSE, align = c("l", "c", "c", "l")) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                            full_width = FALSE,
                            position = "left")
```

</details>

## Pipeline Log

```{r log_warnings, results='asis'}
if (!is.null(params$log_file) && !identical(params$log_file, NA) && file.exists(params$log_file)) {
  log_lines <- readLines(params$log_file, warn = FALSE)

  # Split into warnings and non-warning lines. Drop noisy multi-line Bioconductor seqlevel warnings.
  seqlevel_start <- grepl("sequence levels not in the other", log_lines, fixed = TRUE)
  seqlevel_cont <- grepl("^  - in ", log_lines) | grepl("^  Make sure", log_lines) | grepl("^  genome ", log_lines)
  seqlevel_warning <- seqlevel_start | seqlevel_cont
  warning_lines <- log_lines[startsWith(log_lines, "[WARNING]") & !seqlevel_warning]
  session_lines <- log_lines[grepl("^(R version|Platform|  )", log_lines)]
  info_lines    <- log_lines[!startsWith(log_lines, "[WARNING]") &
                              !grepl("^(R version|Platform|  )", log_lines) &
                              !seqlevel_warning]
  # --- Warnings block ---
  cat('<details>\n<summary><strong>&#9888; Warnings (', length(warning_lines), ')</strong></summary>\n\n', sep = "")
  if (length(warning_lines) > 0) {
    cat('<pre style="max-height:400px;overflow-y:auto;background:#fff8e1;border-left:4px solid #f9a825;padding:8px;">\n')
    cat(paste(htmltools::htmlEscape(warning_lines), collapse = "\n"))
    cat('\n</pre>\n')
  } else {
    cat('<p><em>No warnings recorded.</em></p>\n')
  }
  cat('</details>\n\n')

  # --- Full log block ---
  cat('<details>\n<summary><strong>&#128196; Full pipeline log</strong></summary>\n\n')
  cat('<pre style="max-height:500px;overflow-y:auto;background:#f5f5f5;padding:8px;">\n')
  cat(paste(htmltools::htmlEscape(info_lines), collapse = "\n"))
  cat('\n</pre>\n')
  cat('</details>\n\n')

} else {
  cat("*Log file not available.*\n")
}
```

## Session Info

```{r session_info, results='asis'}
si <- sessionInfo()

si_lines <- capture.output(print(si))

cat('<details>\n<summary><strong>&#128196; Session information</strong></summary>\n\n')
cat('<pre style="max-height:400px;overflow-y:auto;background:#f5f5f5;padding:8px;">\n')
cat(paste(htmltools::htmlEscape(si_lines), collapse = "\n"))
cat('\n</pre>\n')
cat('</details>\n')
```